CE888 Assignment 1

### 2019-02-20 paper/main.tex
update

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{U}{nsupervised} learning is a branch of machine learning that deals with unlabeled data to find patterns, groups
and relations. This sub section of machine learning is useful to describe and understand unknown data, which 
also helps us understand the properties of future data samples. Right now we are collecting data much faster than we can process it.
And most of the time the data we acquire is unlabeled, and these huge amounts of data are complex, noisy
and it might include unnecessary information.\par There are different ways to reduce the complexity of a given data and learn good 
features from it. In this work, we will look at auto encoders, one of those procedures used to learn features from a given data.
Auto encoders are a part of representational learning, which can be trained with or without labels.
They play a key role in deep learning applications. They are used in dimensionality reduction and feature 
extraction, and together with deep learning they gained a lot of traction after Hinton proposed a method to train 
deep belief networks one layer at a time [2]. The motivation for using auto encoders is, in what they do they are
very similar to traditional feature selection algorithms, but they are also capable of learning good features from
complex and high dimensional data. This aspect of auto encoders and unsupervised learning in general is important.
As LeCun states in [3], people are not paying attention to unsupervised learning because of the success of the supervised learning
in recent years. But it will be far more important in the future because of how living beings perceive and learn, they do not
need labels to recognize the world around them.\par In this paper we are going to look at previous work (Section II),
look at different methods and present their results to show how they compare. In Section III, we are going to present 
our goals and what we want to accomplish. Then we will talk about the datasets that we are going to use for the task and 
give detailed information about them. The planned experiments and analyses for the task will be mentioned in Section IV.
Evaluation methods that are going to be used after the experiments will be presented in Section V.

\section{Background}
The conclusion goes here.
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

@@

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{ {./images/} }
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else

  \usepackage[dvips]{graphicx}
  \graphicspath{{../eps/}}
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{U}{nsupervised} learning is a branch of machine learning that deals with unlabeled data to find patterns, groups
and relations. This sub section of machine learning is useful to describe and understand unknown data, which 
also helps us understand the properties of future data samples. Right now we are collecting data much faster than we can process it.
And most of the time the data we acquire is unlabeled, and these huge amounts of data are complex, noisy
and it might include unnecessary information.\par There are different ways to reduce the complexity of a given data and learn good 
features from it. In this work, we will look at auto encoders, one of those procedures used to learn features from a given data.
Auto encoders are a part of representational learning, which can be trained with or without labels.
They play a key role in deep learning applications. They are used in dimensionality reduction and feature 
extraction, and together with deep learning they gained a lot of traction after Hinton proposed a method to train 
deep belief networks one layer at a time [2].Another addition to the auto encoder will be the softmax activation function.
Since softmax gives us the probability distribution over-the classes of our data- the output vector [3], we can use the output of our
encoder to train a clustering algorithm. The motivation for using auto encoders is, in what they do they are
very similar to traditional feature selection algorithms, but they are also capable of learning good features from
complex and high dimensional data. This aspect of auto encoders and unsupervised learning in general is important. 
As LeCun states in [3], people are not paying attention to unsupervised learning because of the success of the supervised learning
in recent years. But it will be far more important in the future because of how living beings perceive and learn, they do not
need labels to recognize the world around them.\par In this paper we are going to look at previous work (Section II),
look at different methods and present their results to show how they compare. In Section III, we are going to present 
our goals and what we want to accomplish. Then we will talk about the datasets that we are going to use for the task and 
give detailed information about them. The planned experiments and analyses for the task will be mentioned in Section IV.
Evaluation methods that are going to be used after the experiments will be presented in Section V.

\section{Background}

\section{Methodology}
Our main goal in this study is to correctly cluster data using unsupervised machine learning techniques.
To accomplish this task we will go through three different subtasks. First, we will use traditional clustering algorithms 
to see how they perform on our datasets. In the second part, we will use auto encoders to learn good features
from our datasets, then we will pass the- extracted features- outputs of the auto encoder to a clustering algorithm to see the difference
in their performance. Finally, we will change the output activation function of the auto encoder to the softmax 
activation function, which will give us the probability distribution of the output vector. Then we will use
the maximum weighted class as our cluster assignment.\par
Before starting the experiments the datasets will be examined and processed appropriately for each sub-task.
For the first sub task, depending on the complexity and dimensions of the given data, a suitable clustering algorithm will be chosen.
For the second sub-task, features of the datasets will be standardized. As LeCun concludes in [3], networks
can learn better and faster if the inputs are centered around zero with equalized covariance. Another consideration for sub-tasks
two and three is using ensemble of auto encoders. Since the optimization of the randomly initialized weights is 
difficult [4], we can have multiple auto encoders in our model to mitigate the effects of random weight initialization.
For the last sub-task, since we will be using softmax as our activation function, the targets of our datasets will be 
converted to one hot coded versions.\par

The three datasets chosen to be clustered in this study are Modified NIST, Human Activity Recognition Using Smartphones (HAR) 
and High Time Resolution Universe Survey 2 (HTRU2).\par

MNIST is a image dataset of hand written digits. It is a modified version of the NIST dataset collected from Census Bureau employees and high-school
students and processed by LeCun et.al..The original NIST set was normalized, so that each sample would be centered
according to the weights of it's pixels in a 28x28 box. Images were also converted to grey-scale and shuffled to make it
easy to use in machine learning applications [5]. The dataset consists of 60.000 training images and 10.000 test images. Since each
sample is represented as a 28x28 pixel binary image, it can be said that samples have 784 features. MNIST dataset is frequently
used in machine learning research[--], which makes it easier for researchers to compare their models and solutions to the others
in the field.\newpage

\begin{figure}[!ht]
\centering
\includegraphics[width=2.5in]{images/cor_heat_mnist.png}
\caption{Corraletion heatmap of MNIST features.}
\label{fig_mnist}
\end{figure}

HAR is a dataset of multiple classes of human activities. It was created by recording different classes of activities of 30 volunteers
by using the accelerometer and the gyroscope in a smartphone. These sensors were used to record the linear acceleration and angular velocity
on three axes at 50 Hz. Volunteers were also recorded by a camera for the labeling of the activities(walking, walking upstairs, walking downstairs, sitting, standing, laying)[7]. The dataset consists of 7352
training samples and 2947 test samples. Each sample has 561 features which are- extracted- estimated from the recorded accelerometer and gyroscope signals.These features are normalized, scaled and shifted to be within [-1,1].  Subject id labels for the features and mentioned raw accelerometer and gyroscope signals are also provided with the dataset.
Dimensions of the subject ids are the same with their training and testing parts. On the other hand, the raw signals are seperated into nine different files grouped under three axes, which are body acceleration (X,Y,Z), body gyroscope (X,Y,Z) and total acceleratin (X,Y,Z).

\begin{figure}[!ht]
  \centering
  \includegraphics[width=2.5in]{images/cor_heat_har.png}
  \caption{Corraletion heatmap of HAR features.}
  \label{fig_har}
  \end{figure}
  \par

  Our last dataset HTRU2 was obtained by Bates[] using ANNs to first filter out the bad candidates,then manually inspecting the outputs of the ANN.Each sample of HTRU2 is a set of features extracted from the data acquired by large radio telescopes during High Time Resolution Universe Survey. HTRU2 has 8 features and consists of 17,898 total samples. Only 1,639 of them are positive pulsar samples and the 16,259 of them are negative samples. The first four features are simple statistics of the pulsar profiles and rest of the features were obtained from DM-SNR curve [].These compact features were obtained from 90.000 labeled pulsar candidates by using Pulsar Feature Lab [].

  \begin{figure}[!ht]
    \centering
    \includegraphics[width=2.5in]{images/cor_heat_htru.png}
    \caption{Corraletion heatmap of htru features.}
    \label{fig_htru}
    \end{figure}
    \par

\section{Experiments}

In the experiments the implementation of the tasks and data analysis will be done using python. In the imlementation of the traditional clustering algorithms, sckit-learn library will be used because of its easy to use nature and rich functions. For the rest of the tasks, tensorflow and keras libraries will be used to implement auto encoders, Since tensorflow converts the python code into C code in runtime, it is fast and it can run on GPUs, making it one of the best neural network library. Finally, the data analysis and plotting will be done using matplotlib and seaborn libraries.

\subsection{Clustering Algorithms}
In the core part of the clustering experiments, k-Means and DBSCAN algorithms will be used. These algorithms will be first compared to each other, then their results will be analysed, evaluated and recorded for future comparisons. If there is enough time to deviate from the core path,agglomerative clustering algorithm will also be used and these clustering algorithms will be tested again after applying dimensionality reduction methods like PCA/kernel PCA, LDA to the datasets. Otherwise, the experiments will continue with the auto encoders.

@@

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.

% \begin{figure*}[!t]
% \centering
% \subfloat[Case I]{\includegraphics[width=2.5in]{box}%
% \label{fig_first_case}}
% \hfil
% \subfloat[Case II]{\includegraphics[width=2.5in]{box}%
% \label{fig_second_case}}
% \caption{Simulation results for the network.}
% \label{fig_sim}
% \end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}

\bibliography{IEEEabrv,../bib/paper}
%
